---
title: "GB demand flexibility service savingSessions"
subtitle: "Analysis of GB electricity demand and twitter data"
author: "Ben Anderson (@dataknut)"
date: 'Last run at: `r Sys.time()`'
format: 
  html:
    self-contained: true
    toc: true
    toc-depth: 3
    code-fold: true
    number-sections: true
execute:
  echo: false
  warning: false
editor: visual
---

# Introduction

UK 'demand flexibility service' [DFS](https://www.nationalgrideso.com/industry-information/balancing-services/demand-flexibility) experiments/trials/'live' events by NG-ESO and retailers such as [\@OctopusEnergy](https://twitter.com/SavingSessions)

 * DFS began on 1 November 2022. 
 * The service will run until March 2023.
 
Relevant media articles:
 
  * Monday 23rd Jan 2023: [https://www.bbc.co.uk/news/business-64367504](https://www.bbc.co.uk/news/business-64367504)
  * Tuesday 24th Jan 2023: [https://www.bbc.co.uk/news/business-64372264](https://www.bbc.co.uk/news/business-64372264)

Questions:

> Q1: Is there a noticeable 'dip' in demand? (Can we even tell?)

> Q2: Do tweets tell us anything useful about the events?

# Code setup

Part of <https://github.com/dataknut/savingSessions>

```{r}
#| label: codeSetup
#| warning: false

library(dkUtils)

myLibs <- c("data.table",
            "dplyr",
            "flextable",
            "ggplot2",
            "hashTagR",
            "hms",
            "knitr",
            "kableExtra",
            "lubridate",
            "readr",
            "rtweet",
            "tidytext",
            "wordcloud")

# load the libraries, install if we don't have them
print("Loading libraries...")
dkUtils::loadLibraries(myLibs) # get it here: devtools::install_github("dataknut/dkUtils")

esoDataPath <- path.expand("~/Dropbox/data/UK_NGESO/genMix/")
twitterDataPath <- path.expand("~/Dropbox/data/twitter/")
  
# functions ----

makeFlexTable <- function(df, cap = "caption", digits = 0){
  # makes a pretty flextable - see https://cran.r-project.org/web/packages/flextable/index.html
  ft <- flextable::flextable(df)
  ft <- colformat_double(ft, digits = digits)
  ft <- fontsize(ft, size = 9)
  ft <- fontsize(ft, size = 10, part = "header")
  ft <- set_caption(ft, caption = cap)
  return(flextable::autofit(ft))
}

# generic plotting function - choose the variable you want to compare and amend the y axis label to match
make_comparisonPlot <- function(dt, # half hourly data
                                startDateTime,
                                endDateTime,
                                yVar = "GENERATION", #default
                                yVarLab = "MW per half-hour",
                                timeLag = 10 # default number of comparison days
                                  ){
  # for testing
  # dt <- ngeso_dt_orig
  # startDateTime <- session1Start
  # endDateTime <- session1End
  # yVar <- "GENERATION"
  # yVarLab = "MW per half-hour"
  # timeLag = 10 # default number of comparison days
  
  
  res <- list() # results holder
  dt[, dv_hms := hms::as_hms(dv_start)]
  
  sessionDate <- lubridate::date(startDateTime) # you'll see why
  if(lubridate::wday(startDateTime) < 6){
    sessionDay <- "weekday" # you'll see why
  } else {
    sessionDay <- "weekend"
  }
  # 1 = Sunday, 7 = Saturday etc
  dt[, ba_wday := lubridate::wday(dv_start, label = TRUE)]
  dt[, ba_wd := ifelse(lubridate::wday(dv_start) > 1 & lubridate::wday(dv_start) < 7,
                       "weekday",
                       "weekend")]
  # extract and aggregate baseline
  # first get similar days (weekdays or weekends as needed)
  
  similarDays <- dt[dv_date != sessionDate & ba_wd == sessionDay &
                      dv_date < sessionDate]
  uniqueN(similarDays$dv_date)
  dates <- tail(similarDays[,(nObs = .N), keyby = .(dv_date)], timeLag) # get the most recent n = timeLag days
  datesToGet <- dates[, dv_date]
  daysWeWant <- similarDays[dv_date %in% datesToGet]
  uniqueN(daysWeWant$dv_date)
  
  # then get the n = timeLag most recent
  baseline_HalfHourlyDT <- daysWeWant[,
                              .(y_mean = mean(get(yVar))), # mean of the chosen variable
                              keyby = .(hms = dv_hms, 
                                        ba_wd) # to check
                              ]
  #table(baseline_HalfHourlyDT$ba_wday)
  res$nCompDays <- uniqueN(daysWeWant$dv_date) # just to be sure
  baseline_HalfHourlyDT[, legend_lab := paste0("Comparison mean")]
  # extract session day
  session_HalfHourlyDT <- dt[dv_date == sessionDate, # the session day,
                              .(y_mean = mean(get(yVar))), # mean of just one value but this is just to match baseline variable names
                              keyby = .(hms = dv_hms,
                                        ba_wd)
                              ]
  session_HalfHourlyDT[, legend_lab := "Saving session"]
  
  # combine them in long form for ggplot
  plotDT <- rbind(baseline_HalfHourlyDT, session_HalfHourlyDT)
  table(plotDT$ba_wd)
  plotDT[, adjusted_hms := hms::as_hms(hms + (15*60))] # so they plot at centres of half-hours
  
  periodAlpha <- 0.3 #  shaded rects on plots
  periodFill <- "grey50"
    ymax <- max(plotDT$y_mean) # GW
    ymin <- min(plotDT$y_mean) # GW
    xmin <- hms::as_hms(lubridate::as_datetime(startDateTime))
    xmax <- hms::as_hms(lubridate::as_datetime(endDateTime)) + 30*60 # to allow for the start time
    
    t <- plotDT[hms >= xmin & hms < xmax]
    
    # be explicit about using data.table::dcast
    wt <- data.table::dcast(t[, .(hms, wday, y_mean, legend_lab)], hms ~ legend_lab, value.var = "y_mean")
    
    wt[, diff := `Saving session` - `Comparison mean`]
    res$wt <- wt[, pc_diff := 100*(diff/`Comparison mean`)]
    
    label <- paste0("SavingSession (", 
                    as.Date(startDateTime),
                    ")"
    )
    plotDT[, legend_lab := ifelse(legend_lab == "Saving session",
                                  label,
                                  legend_lab)]
    
    res$p <- ggplot2::ggplot(plotDT, aes(x = adjusted_hms, y = y_mean, 
                                         colour = legend_lab)) +
      geom_line() +
      geom_point() +
      annotate("rect", xmin = xmin,
               xmax = xmax,
               ymin = ymin, ymax = ymax,
               alpha = periodAlpha, fill = periodFill) +
      scale_color_manual(name = "Legend", values=c('grey', 'red')) + # this should always make comparison grey
      theme(legend.position = "bottom") +
      labs(x = "Time of day",
           y = yVarLab,
           caption = paste0("Saving session: ", pdt(startDateTime)," to ", pdt(endDateTime + 30*60),
                            "\nData: NG-ESO GB Generation Mix (y variable: ", yVar ,")\nPlot: @dataknut, points plotted at middle of half-hour for clarity\nNumber of comparison days: ", res$nCompDays)
      )
    
    return(res)
}


pdt <- function(t){ # make a pretty date
  strftime(t, "%a %d %b %Y %H:%M")
}

pd <- function(t){ # make a pretty date
  strftime(t, "%a %d %b %Y")
}

# parameters ----
# > load session dates from file ----

savingSessions_dateTimes_DT <- data.table::as.data.table(readxl::read_xlsx(here::here("data", "savingSessions_dateTimes.xlsx")))

```


# Data

## Electricity demand and carbon intensity data

Use the half-hourly NG-ESO 'generation mix' [data](https://data.nationalgrideso.com/carbon-intensity1/historic-generation-mix) as a reasonable proxy for 'demand'. This also includes a useful estimate of carbon intensity.

> Should probably use the [demand data](https://data.nationalgrideso.com/demand/historic-demand-data) instead as it includes embedded wind & solar.

```{r}
#| label: getDemandData
#| warning: false

esoF <- paste0(esoDataPath, "latest_df_fuel_ckan.csv")
# check for gzipped version (see below)
if(file.exists(paste0(esoF, ".gz"))){ 
  message("We already have a version saved to: ", paste0(esoF, ".gz"))
  message("Loading it...")
  ngeso_dt_orig <- data.table::fread(paste0(esoF, ".gz"))
} else {
  message("We don't already have a version, downloading and saving to: ", esoF)
  ngeso_dt_orig <- data.table::fread("https://data.nationalgrideso.com/backend/dataset/88313ae5-94e4-4ddc-a790-593554d8c6b9/resource/f93d1835-75bc-43e5-84ad-12472b180a98/download/df_fuel_ckan.csv")
  # nice dateTime
  ngeso_dt_orig[, dv_start := lubridate::as_datetime(DATETIME)]
  data.table::fwrite(ngeso_dt_orig, esoF) # save locally for future re-use
  dkUtils::gzipIt(esoF)
}

# if older than 1 day, reload
today <- lubridate::today()
lastNGESO <- as.Date(max(ngeso_dt_orig$dv_start))

if(today - lastNGESO >= 1) {
  # old data, reload
  message("But the version we have dates from ", lastNGESO, " (",today - lastNGESO ," days ago), downloading latest...")
  ngeso_dt_orig <- data.table::fread("https://data.nationalgrideso.com/backend/dataset/88313ae5-94e4-4ddc-a790-593554d8c6b9/resource/f93d1835-75bc-43e5-84ad-12472b180a98/download/df_fuel_ckan.csv")
  # nice dateTime
  ngeso_dt_orig[, dv_start := lubridate::as_datetime(DATETIME)]
  data.table::fwrite(ngeso_dt_orig, esoF)
  dkUtils::gzipIt(esoF)
}

# we think renewable is wind + solar, low carbon includes nuclear

ngeso_dt_orig[, dv_date := lubridate::as_date(DATETIME)] # for filtering etc
ngeso_dt_orig[, dv_hms := hms::as_hms(DATETIME)] # for filtering etc
```

Variables of interest:

 * GENERATION is in MW. It has some exclusions, its just an indicator.
 * CARBON_INTENSITY is in g CO2/kWh - we'd expect this to be high during *DFS* events (as we're trying to avoid the use of high carbon & expensive generation?)

## Twitter data

Use [rtweet](https://docs.ropensci.org/rtweet/) to collect tweets matching:

 * `#savingSession` or `#savingSessions` (not case sensitive) - this seems to be the preferred [Octopus hashtag](https://twitter.com/hashtag/savingSession)
 * `#DemandFlexibilityService` (not case sensitive) - as used by [NG-ESO](https://twitter.com/hashtag/DemandFlexibilityService), appears less frequently used

We keep these datasets distinct as not all Octopus DFS events were also wider NG-ESO 'live' events.

We do NOT store the tweets in the repo for both [ethical](https://blogs.lse.ac.uk/impactofsocialsciences/2015/09/28/challenges-of-using-twitter-as-a-data-source-resources/) and practical reasons...

Note also that we may not be collecting the complete dataset of hashtagged tweets due to:

 * the [intricacies of the twitter API](https://www.demos.co.uk/files/Road_to_representivity_final.pdf?1441811336).
 * [changes to the twitter API policy](https://twitter.com/TwitterDev) which may have reduced the availability of tweets from February 2023. It is currently unclear whether a [free API access tier](https://developer.twitter.com/en/products/twitter-api) will be available or [per-project academic research access](https://developer.twitter.com/en/products/twitter-api/academic-research) will be required.

It is important to note that both datasets are therefore likely to be partial.

### #savingSessions tweets

@fig-getSavingSessionTweets shows the cumulative number of #savingSession tweets collected and marks the relevant events. Labels are likely to overlap where events are close together.

```{r}
#| label: fig-getSavingSessionTweets
#| fig-cap: "Cumulative number of #savingSession tweets collected"
#| warning: FALSE

hashtags <- c("savingSession", "savingSessions", "savingsession") # fairly sure it's case insensitive
searchString <- hashTagR::createSearchFromTags(hashtags) # convert to rtweet search string

now <- lubridate::now()

dataPath <- path.expand(paste0(twitterDataPath, "savingSessions/"))

if(dir.exists(dataPath)){
  ofile <- path.expand(paste0(dataPath, "tw_",now,".csv"))
  tweetsDT <- hashTagR::getTweets(searchString, n = 20000)
  if(nrow(tweetsDT)>0){
    # only write out a file if there are new tweets to avoid confusing hashTagR::loadTweets later on
    readr::write_csv(tweetsDT, file = ofile) # data.table breaks here
  #message("Retrieved ", nrow(tweetsDT), " tweets and saved them to ", ofile)
  } else {
    message("No recent tweets found using ", searchString, " - loading previous tweets")
  }
  #message("Retrieved ", nrow(tweetsDT), " tweets and saved them to ", ofile)
} else {
  message(dataPath, " not found!")
}

# now load all the tweet files we have matching that searchString

tweetsDT <- hashTagR::loadTweets(path = dataPath,
                                 pattern = "*.csv",
                                 noisy = FALSE)

# clean up the tweet text to remove stuff we don't want.
cleanTweets <- function(dt){
  # expects a data.table
  dt[, clean_text := gsub("https\\S*", " ", text)]
  dt[, clean_text := gsub("@\\S*", " ", clean_text)]
  dt[, clean_text := gsub("amp", " ", clean_text)]
  dt[, clean_text := gsub("RT", " ", clean_text)]
  dt[, clean_text := gsub("[\r\n]", " ", clean_text)]
  dt[, clean_text := gsub("[[:punct:]]", " ", clean_text)]
  return(dt)
}

# clean up the tweet text to remove stuff we don't want.
uniq_savingSessionTweetsDT <- cleanTweets(hashTagR::processTweets(tweetsDT))

plotDT <- uniq_savingSessionTweetsDT[, .(nObs = .N), keyby = .(Time = ba_created_at_dateHour)]

plotDT[, cumSum := cumsum(nObs)]

# plot over time

myCaption <- "Data: twitter API via @mkearney's rtweet package\nPlot by @dataknut\nHashtag: #savingSession\nOctopus #savingSessions marked"
p <- ggplot2::ggplot(plotDT, aes(x = Time, y = cumSum)) + 
  geom_point() +
  labs(y = "Cumulative number of tweets",
       caption = myCaption)

add_session_label <- function(p, startDateTime, endDatetime){
  p <- p +
    geom_vline(xintercept = startDateTime, alpha = periodAlpha, colour = periodFill) +
    annotate("text", x = startDateTime, vjust=-0.1, angle=90,
             y = ymax*0.5, label = pd(startDateTime))
  return(p)
}

add_all_session_labels <- function(p){
  for(s in savingSessions_dateTimes_DT$ID) {
    sessionStart <- savingSessions_dateTimes_DT[ID == s, start_hh]
    #print(sessionStart)
    sessionEnd <- savingSessions_dateTimes_DT[ID == s, end_hh]
    #print(sessionEnd)
    p <- add_session_label(p, sessionStart, sessionEnd)
  }
  return(p)
}

periodAlpha <- 0.3 #  shaded rects on plots
periodFill <- "red"
ymax <- max(plotDT$cumSum)
ymin <- min(plotDT$cumSum)

add_all_session_labels(p)
```


```{r}

ssTweets <- uniq_savingSessionTweetsDT # all of them
```

Number of tweets: `r nrow(ssTweets)`

*Positive words...*

```{r}
#| warning: true

# default cloud = positive
makeSentimentCloud <- function(dt, sentiFilter = "positive"){
  # start with the selected tweets
  # grab the words
  tweets_words <-  dt %>%
    dplyr::select(clean_text) %>%
    tidytext::unnest_tokens(word, clean_text)
  # count the words
  words <- tweets_words %>% count(word, sort=TRUE)
  # remove the stop words
  data(stop_words)
  
  # remove the stop words - they are usually the most frequent
  # and usually the least interesting
  de_stopped <- words %>%
    dplyr::anti_join(stop_words)
  
  bsc <- de_stopped %>%
    dplyr::inner_join(tidytext::get_sentiments("bing")) %>% # only keeps those with sentiments
    #count(linenumber, sentiment, sort = TRUE) %>%
    ungroup()
  
  results <- list()
  results$ft <- table(bsc$sentiment)
  
  results$nt <- bsc %>%
    group_by(sentiment) %>%
    summarize(freq=sum(n))
  
  # select the sentiment to wordCloud
  filtered <- bsc %>%
    dplyr::filter(sentiment == sentiFilter) # why not working?
  
  set.seed(1234) # for reproducibility
  # seems to get rendered immediately, why?
  wordcloud::wordcloud(words = filtered$word, 
                             freq = filtered$n, min.freq = 1,           
                             max.words=200, 
                             random.order=FALSE, 
                             rot.per=0.35,            
                             colors=brewer.pal(8, "Dark2"))
  results$dates <- paste0(pdt(min(dt$ba_created_at_date)),
                      " to ", pdt(max(dt$ba_created_at_date)))
  return(results)
}


# renders plot immediately (not sure why)
res <- makeSentimentCloud(ssTweets, 
                   sentiFilter = "positive")

message("\nTweets from ", res$dates)
```

*Negative words...*

```{r}
#| warning: true

res <- makeSentimentCloud(ssTweets, 
                   sentiFilter = "negative")
```

Total word and unique word counts...

```{r}
knitr::kable(res$ft, 
             caption = paste0("Number of unique words by sentiment")
             ) %>%
  kable_styling()

knitr::kable(res$nt, 
             caption = paste0("Number of words by sentiment")
             ) %>%
  kable_styling()

```

### #demandFlexibilityService tweets

```{r}

hashtags <- c("DemandFlexibilityService") # fairly sure it's case insensitive
searchString <- hashTagR::createSearchFromTags(hashtags) # convert to rtweet search string

now <- lubridate::now()

dataPath <- path.expand(paste0(twitterDataPath, "demandFlexibilityService/"))

if(dir.exists(dataPath)){
  ofile <- path.expand(paste0(dataPath, "tw_",now,".csv"))
  tweetsDT <- hashTagR::getTweets(searchString, n = 20000)
  if(nrow(tweetsDT)>0){
    # only write out a file if there are new tweets to avoid confusing hashTagR::loadTweets later on
    readr::write_csv(tweetsDT, file = ofile) # data.table breaks here
  #message("Retrieved ", nrow(tweetsDT), " tweets and saved them to ", ofile)
  } else {
    message("No recent tweets found using ", searchString, " - loading previous tweets")
  }
} else {
  message(dataPath, " not found!")
}

# now load all the tweet files we have matching that searchString

tweetsDT <- hashTagR::loadTweets(path = dataPath, 
                                 pattern = "*.csv",
                                 noisy = TRUE)

uniq_demandFlexibilityServiceTweetsDT <- cleanTweets(hashTagR::processTweets(tweetsDT))

dfsTweets <- uniq_demandFlexibilityServiceTweetsDT # all
```

Number of tweets: `r nrow(dfsTweets)`

*Positive words...*

```{r}
#| warning: true

# renders plot immediately (not sure why)
res <- makeSentimentCloud(dfsTweets, 
                   sentiFilter = "positive")
message("\nTweets from ", res$dates)
```

*Negative words...*

```{r}
#| warning: true

res <- makeSentimentCloud(dfsTweets, 
                   sentiFilter = "negative")
```

Total word and unique word counts...

```{r}
knitr::kable(res$ft, 
             caption = paste0("Number of unique words by sentiment")
             ) %>%
  kable_styling()

knitr::kable(res$nt, 
             caption = paste0("Number of words by sentiment")
             ) %>%
  kable_styling()

```



@fig-getDemandFlexibilityServiceTweets shows the cumulative number of #DemandFlexibilityService tweets collected and marks the relevant events. Labels are likely to overlap where events are close together.

> NB: data collection for this hashtag only started on 23rd January - it looks like twitter's search API does not return tweets more than x days old. People were tweeting this hashtag [earlier than January](https://twitter.com/hashtag/DemandFlexibilityService?src=hashtag_click&f=live)...

```{r}
#| label: fig-getDemandFlexibilityServiceTweets
#| fig-cap: "Cumulative number of #DemandFlexibilityService tweets collected"
#| warning: FALSE

plotDT <- uniq_demandFlexibilityServiceTweetsDT[, .(nObs = .N), 
                                                keyby = .(Time = ba_created_at_dateHour)]

plotDT[, cumSum := cumsum(nObs)]

# plot over time

myCaption <- "Data: twitter API via @mkearney's rtweet package\nPlot by @dataknut\nHashtag: #demandFlexibilityService\nNG-ESO 'live' DFS events marked"
p <- ggplot2::ggplot(plotDT, aes(x = Time, y = cumSum)) + 
  geom_point() +
  labs(y = "Cumulative number of tweets",
       caption = myCaption)

add_ngeso_session_labels <- function(p){
  sessions <- savingSessions_dateTimes_DT[type == "NG-ESO live", ID]
  for(s in sessions) {
    sessionStart <- savingSessions_dateTimes_DT[ID == s, start_hh]
    #print(sessionStart)
    sessionEnd <- savingSessions_dateTimes_DT[ID == s, end_hh]
    #print(sessionEnd)
    p <- add_session_label(p, sessionStart, sessionEnd)
  }
  return(p)
}

periodAlpha <- 0.3 #  shaded rects on plots
periodFill <- "red"
ymax <- max(plotDT$cumSum)
ymin <- min(plotDT$cumSum)
add_ngeso_session_labels(p)
```

# Analysis

Q1: Demand reduction

Simple comparison of the 'demand response session' day half-hourly demand with the mean half-hourly demand for the previous n weekdays or weekend days as appropriate. Default n is 10.

All sorts of caveats apply:

 * not all of these sessions were national 'live' events, some were trials by a single retailer (Octopus)
 * even though we compare with weekdays or weekend days as appropriate, we may still not be comparing like with like (cold vs warm, holiday vs non-holiday etc)
 * we can't tell from this analysis what demand would have looked like _without_ the demand response intervention. That would require all the retailers to aggregate the 'savings' they estimate... or a randomised control trial

Q2: Twitter sentiment analysis

Inspired by <https://www.tidytextmining.com/sentiment.html> (sentiment analysis) and <https://towardsdatascience.com/create-a-word-cloud-with-r-bde3e7422e8a> (word clouds)

Extract the tweets for the day of each *DFS* event _and the day after_.

Remove stop words (to, the, and, a, for, etc) & profanity etc from the tweets and sentiment them.

In each case we report the number of negative and positive codings (according to `tidytext::get_sentiments("bing")`) for the unique words. This will add up to the Number of unique words by sentiment. We then report the total frequency of words that are negative or positive (which will add up to the total number of words).

Remember the size of the (positive/negative) words in the word cloud is relative to the count of all (positive/negative) words in the wordcloud, not the total word count.


```{r}
#| label: sessionDates
#| tbl-cap: "Saving sessions"

makeFlexTable(savingSessions_dateTimes_DT,cap = "default caption")
```


## Session 1 (Octopus: `r pdt(savingSessions_dateTimes_DT[ID == 1, start_hh])`)

```{r}
start_hh <- savingSessions_dateTimes_DT[ID == 1, start_hh]
end_hh <- savingSessions_dateTimes_DT[ID == 1, end_hh]
actualEnd <- savingSessions_dateTimes_DT[ID == 1, end_hh] + 30*60
```

 * `r start_hh` to
 * `r actualEnd`
 
(Not all retailers may have taken part in this one)

According to <https://energycentral.com/news/octopus-energy-provide-108mw-grid-flexibility-first-%E2%80%98saving-session%E2%80%99>:

> "Over 200,000 households reduced their energy demand by 108MW collectively, the same as a gas power station can generate in an hour. If replicated by all UK energy suppliers, this would be over 1GW."

But <https://www.current-news.co.uk/news/octopus-energy-customers-provide-108mw-of-flexibility-during-first-saving-session> quotes Octopus as actually saying:

> "if the program was scaled to _all electric smart meter customers_ with all supplier\[s\] in Great Britain, it could create over 1GW of flexible energy load"

The following plot compares GB electricity demand during the *DFS* event with mean GB electricity demand over the last n similar days (i.e. weekdays or weekends as appropriate).

```{r}
#| label: makePlotSession1
#| tbl-cap: "MW comparisons"
#| warning: false
session1_dt <- ngeso_dt_orig[dv_start >= start_hh & 
                               dv_start < end_hh]

session1_mean_mw <- mean(session1_dt$GENERATION)
session1_mean_gw <- session1_mean_mw/1000

scaled_kW <- (108/200)*28000000
scaled_GW <- ((108/200)*28000000)/1000000

scaled_GW_pc <- 100*(scaled_GW/session1_mean_gw)

gb_smartMeters <- 14000000
mean_kW_per_participant <- 108/200
scaled_GB_smart_meters_GW_pc <- 100*((gb_smartMeters*mean_kW_per_participant)/1000000)/session1_mean_gw

res <- make_comparisonPlot(ngeso_dt_orig, start_hh, end_hh)
#res$p # the plot
res$p

makeFlexTable(res$wt[, hms:= as.factor(hms)], 
              cap = paste0("MW comparisons (hms = half hour period start, number of comparison days = ", res$nCompDays,")"),
              digits = 1) # the table
```

> Confounding factors:

 * recent temperature trends
 * recent holidays
 * etc

>Note that the comparison plot/table shows the difference between the *DFS* event day and the previous similar days. This does not show what the event day energy use would have been in the absence of the intervention - we'd need a randomised control on the day for that (or a sophisticated demand model as the counterfactual).

Octopus customers' 108 MW was `r round(100*(108/session1_mean_mw),3)` % of mean GB generation (`r session1_mean_gw` GW) over the hour of the session (see NG-ESO data above).

108 MW over 200,000 customers is `r 108/200000` MW per household or, more sensibly, `r mean_kW_per_participant` kW per household. A pretty reasonable reduction.

But if 200,000 produced a 108 MW power reduction (`r mean_kW_per_participant` kW per household ), the 1 GW estimate implies we only have `r prettyNum(1000000/mean_kW_per_participant, big.mark = ",")` GB households with a smart meter. Really? Or did they mean all suppliers taking part in the scheme?

Anyway, the latest data has 14 million domestic eletricity smart meters as of [Q1 2022](https://assets.publishing.service.gov.uk/government/uploads/system/uploads/attachment_data/file/1099629/Q2_2022_Smart_Meters_Statistics_Report.pdf) - 45% of all domestic meters.

So:

-   If all [\~28 million households](https://www.ons.gov.uk/peoplepopulationandcommunity/birthsdeathsandmarriages/families/bulletins/familiesandhouseholds/2021) in the UK did this it would be `r scaled_kW` kW. That's `r scaled_GW` GW...
-   If we reduce ambition a bit and assume that only the 14 million domestic GB customers with a smart meter as of Q2 2022 did this then that would provide `r (gb_smartMeters*mean_kW_per_participant)/1000000` GW. Which would be `r round(scaled_GB_smart_meters_GW_pc,2)` % of generation. Not bad aye?

The trouble is... these are octopus customers self-selecting into a trial. You just have to [read the tweets](https://twitter.com/SavingSessions) to see that there are households with installed batteries, pausable heat pumps in well-insulated homes and EV V2G. At what point will we all be similarly equipped?

### Carbon intensity

And just to keep in mind the carbon context:

```{r}
#| label: fig-firstSavingSessionsCI
#| fig-cap: "Comparison of carbon intensity"
#| warning: false

res <- make_comparisonPlot(ngeso_dt_orig, start_hh, end_hh, 
                           yVar = "CARBON_INTENSITY",
                           yVarLab = "g CO2/kWh")
#res$p # the plot
res$p
```

Turning to hashtags, we only have #savingSessions tweets for this event.

### #savingSessions tweets

```{r}

sessionTweets <- uniq_savingSessionTweetsDT[ba_created_at_date >= as.Date(start_hh) & # on the day
                                 ba_created_at_date <= as.Date(end_hh)+1] # day plus 1
```

Number of tweets: `r nrow(sessionTweets)`

*Positive words...*

```{r}
#| warning: true

# renders plot immediately (not sure why)
res <- makeSentimentCloud(sessionTweets, 
                   sentiFilter = "positive")

message("\nTweets from ", res$dates)
```

*Negative words...*

```{r}
#| warning: true

res <- makeSentimentCloud(sessionTweets, 
                   sentiFilter = "negative")
```

Total word and unique word counts...

```{r}
knitr::kable(res$ft, 
             caption = paste0("Number of unique words by sentiment")
             ) %>%
  kable_styling()

knitr::kable(res$nt, 
             caption = paste0("Number of words by sentiment")
             ) %>%
  kable_styling()

```


## Session 2 (Octopus: `r pdt(savingSessions_dateTimes_DT[ID == 2, start_hh])`)

```{r}
start_hh <- savingSessions_dateTimes_DT[ID == 2, start_hh]
end_hh <- savingSessions_dateTimes_DT[ID == 2, end_hh]
actualEnd <- savingSessions_dateTimes_DT[ID == 2, end_hh] + 30*60
```

 * `r start_hh` to
 * `r actualEnd`
 
(Not all retailers may have taken part in this one)

The same caveats about the comparisons apply here...

```{r}
#| label: makePlotSession2
#| tbl-cap: "MW comparisons"
#| warning: false

res <- make_comparisonPlot(ngeso_dt_orig, start_hh, end_hh)
res$p # the plot

makeFlexTable(res$wt[, hms:= as.factor(hms)], 
              cap = paste0("MW comparisons (hms = half hour period start, number of comparison days = ", res$nCompDays,")"),
              digits = 1) # the table
```

### Carbon intensity

And just to keep in mind the carbon context:

```{r}
#| label: fig-secondSavingSessionsCI
#| fig-cap: "Comparison of carbon intensity"
#| warning: false
res <- make_comparisonPlot(ngeso_dt_orig, start_hh, end_hh, 
                           yVar = "CARBON_INTENSITY",
                           yVarLab = "g CO2/kWh")
res$p

```

Turning to hashtags, we only have #savingSessions tweets for this event.

### #savingSessions tweets

```{r}

sessionTweets <- uniq_savingSessionTweetsDT[ba_created_at_date >= as.Date(start_hh) & # on the day
                                 ba_created_at_date <= as.Date(end_hh+1)] # day plus 1
```

Number of tweets: `r nrow(sessionTweets)`

*Positive words...*

```{r}
#| warning: true


res <- makeSentimentCloud(sessionTweets, 
                   sentiFilter = "positive")

message("\nTweets from ", res$dates)
```

*Negative words...*

```{r}
#| warning: true

res <- makeSentimentCloud(sessionTweets, 
                   sentiFilter = "negative")
```

Total word and unique word counts...

```{r}
knitr::kable(res$ft, 
             caption = paste0("Number of unique words by sentiment")
             ) %>%
  kable_styling()

knitr::kable(res$nt, 
             caption = paste0("Number of words by sentiment")
             ) %>%
  kable_styling()

```

## Session 3 (Octopus: `r pdt(savingSessions_dateTimes_DT[ID == 3, start_hh])`)

```{r}
start_hh <- savingSessions_dateTimes_DT[ID == 3, start_hh]
end_hh <- savingSessions_dateTimes_DT[ID == 3, end_hh]
actualEnd <- savingSessions_dateTimes_DT[ID == 3, end_hh] + 30*60
```

 * `r start_hh` to
 * `r actualEnd`
 
(Not all retailers may have taken part in this one)

The same caveats about the comparisons apply here...

```{r}
#| label: makePlotSession3
#| tbl-cap: "MW comparisons"
#| warning: false

res <- make_comparisonPlot(ngeso_dt_orig, start_hh, end_hh)
res$p # the plot

makeFlexTable(res$wt[, hms:= as.factor(hms)], 
              cap = paste0("MW comparisons (hms = half hour period start, number of comparison days = ", res$nCompDays,")"),
              digits = 1) # the table
```

### Carbon intensity

And just to keep in mind the carbon context:

```{r}
#| label: fig-thirdSavingSessionsCI
#| fig-cap: "Comparison of carbon intensity"
#| warning: false
res <- make_comparisonPlot(ngeso_dt_orig, start_hh, end_hh, 
                           yVar = "CARBON_INTENSITY",
                           yVarLab = "g CO2/kWh")
res$p
```

Turning to hashtags, we only have #savingSessions tweets for this event.

### #savingSessions tweets

```{r}

sessionTweets <- uniq_savingSessionTweetsDT[ba_created_at_date >= as.Date(start_hh) & # on the day
                                 ba_created_at_date <= as.Date(end_hh)+1] # day plus 1
```

Number of tweets: `r nrow(sessionTweets)`

*Positive words...*

```{r}
#| warning: true

# renders plot immediately (not sure why)
res <- makeSentimentCloud(sessionTweets, 
                   sentiFilter = "positive")
message("\nTweets from ", res$dates)
```

*Negative words...*

```{r}
#| warning: true

res <- makeSentimentCloud(sessionTweets, 
                   sentiFilter = "negative")
```

Total word and unique word counts...

```{r}
knitr::kable(res$ft, 
             caption = paste0("Number of unique words by sentiment")
             ) %>%
  kable_styling()

knitr::kable(res$nt, 
             caption = paste0("Number of words by sentiment")
             ) %>%
  kable_styling()

```

## Session 4 (Octopus: `r pdt(savingSessions_dateTimes_DT[ID == 4, start_hh])`)

```{r}
start_hh <- savingSessions_dateTimes_DT[ID == 4, start_hh]
end_hh <- savingSessions_dateTimes_DT[ID == 4, end_hh]
actualEnd <- savingSessions_dateTimes_DT[ID == 4, end_hh] + 30*60
```

 * `r start_hh` to
 * `r actualEnd`
 
(Not all retailers may have taken part in this one)

The same caveats about the comparisons apply here...

```{r}
#| label: makePlotSession4
#| tbl-cap: "MW comparisons"
#| warning: false

res <- make_comparisonPlot(ngeso_dt_orig, start_hh, end_hh)
res$p # the plot

makeFlexTable(res$wt[, hms:= as.factor(hms)], 
              cap = paste0("MW comparisons (hms = half hour period start, number of comparison days = ", res$nCompDays,")"),
              digits = 1) # the table
```

### Carbon intensity

And just to keep in mind the carbon context:

```{r}
#| label: fig-fourthSavingSessionsCI
#| fig-cap: "Comparison of carbon intensity"
#| warning: false
res <- make_comparisonPlot(ngeso_dt_orig, start_hh, end_hh, 
                           yVar = "CARBON_INTENSITY",
                           yVarLab = "g CO2/kWh")
res$p
```

Turning to hashtags, we only have #savingSessions tweets for this event.

### #savingSessions tweets

```{r}

sessionTweets <- uniq_savingSessionTweetsDT[ba_created_at_date >= as.Date(start_hh) & # on the day
                                 ba_created_at_date <= as.Date(end_hh)+1] # day plus 1
```

Number of tweets: `r nrow(sessionTweets)`

*Positive words...*

```{r}
#| warning: true

# renders plot immediately (not sure why)
res <- makeSentimentCloud(sessionTweets, 
                   sentiFilter = "positive")
message("\nTweets from ", res$dates)
```

*Negative words...*

```{r}
#| warning: true

res <- makeSentimentCloud(sessionTweets, 
                   sentiFilter = "negative")
```

Total word and unique word counts...

```{r}
knitr::kable(res$ft, 
             caption = paste0("Number of unique words by sentiment")
             ) %>%
  kable_styling()

knitr::kable(res$nt, 
             caption = paste0("Number of words by sentiment")
             ) %>%
  kable_styling()

```

## Session 5 (Octopus: `r pdt(savingSessions_dateTimes_DT[ID == 5, start_hh])`)

```{r}
start_hh <- savingSessions_dateTimes_DT[ID == 5, start_hh]
end_hh <- savingSessions_dateTimes_DT[ID == 5, end_hh]
actualEnd <- savingSessions_dateTimes_DT[ID == 5, end_hh] + 30*60
```

 * `r start_hh` to
 * `r actualEnd`
 
(Not all retailers may have taken part in this one)

The same caveats about the comparisons apply here...

```{r}
#| label: makePlotSession5
#| tbl-cap: "MW comparisons"
#| warning: false

res <- make_comparisonPlot(ngeso_dt_orig, start_hh, end_hh)
res$p # the plot

makeFlexTable(res$wt[, hms:= as.factor(hms)], 
              cap = paste0("MW comparisons (hms = half hour period start, number of comparison days = ", res$nCompDays,")"),
              digits = 1) # the table
```

### Carbon intensity

And just to keep in mind the carbon context:

```{r}
#| label: fig-fifthSavingSessionsCI
#| fig-cap: "Comparison of carbon intensity"
#| warning: false

res <- make_comparisonPlot(ngeso_dt_orig, start_hh, end_hh, 
                           yVar = "CARBON_INTENSITY",
                           yVarLab = "g CO2/kWh")
res$p

```


Turning to hashtags, we only have #savingSessions tweets for this event.


### #savingSessions tweets

```{r}

sessionTweets <- uniq_savingSessionTweetsDT[ba_created_at_date >= as.Date(start_hh) & # on the day
                                 ba_created_at_date <= as.Date(end_hh)+1] # day plus 1
```

Number of tweets: `r nrow(sessionTweets)`

*Positive words...*

```{r}
#| warning: true

# renders plot immediately (not sure why)
res <- makeSentimentCloud(sessionTweets, 
                   sentiFilter = "positive")
message("\nTweets from ", res$dates)
```

*Negative words...*

```{r}
#| warning: true

res <- makeSentimentCloud(sessionTweets, 
                   sentiFilter = "negative")
```

Total word and unique word counts...

```{r}
knitr::kable(res$ft, 
             caption = paste0("Number of unique words by sentiment")
             ) %>%
  kable_styling()

knitr::kable(res$nt, 
             caption = paste0("Number of words by sentiment")
             ) %>%
  kable_styling()

```

## Session 6 (Octopus: `r pdt(savingSessions_dateTimes_DT[ID == 6, start_hh])`)

```{r}
start_hh <- savingSessions_dateTimes_DT[ID == 6, start_hh]
end_hh <- savingSessions_dateTimes_DT[ID == 6, end_hh]
actualEnd <- savingSessions_dateTimes_DT[ID == 6, end_hh] + 30*60
```

 * `r start_hh` to
 * `r actualEnd`
 
(Not all retailers may have taken part in this one)

The same caveats about the comparisons apply here...

```{r}
#| label: makePlotSession6
#| tbl-cap: "MW comparisons"
#| warning: false

res <- make_comparisonPlot(ngeso_dt_orig, start_hh, end_hh)
res$p # the plot

makeFlexTable(res$wt[, hms:= as.factor(hms)], 
              cap = paste0("MW comparisons (hms = half hour period start, number of comparison days = ", res$nCompDays,")"),
              digits = 1) # the table
```

### Carbon intensity

And just to keep in mind the carbon context:

```{r}
#| label: fig-sixthSavingSessionsCI
#| fig-cap: "Comparison of carbon intensity"
#| warning: false
res <- make_comparisonPlot(ngeso_dt_orig, start_hh, end_hh, 
                           yVar = "CARBON_INTENSITY",
                           yVarLab = "g CO2/kWh")
res$p

```


Turning to hashtags, we only have #savingSessions tweets for this event.

### #savingSessions tweets

```{r}

sessionTweets <- uniq_savingSessionTweetsDT[ba_created_at_date >= as.Date(start_hh) & # on the day
                                 ba_created_at_date <= as.Date(end_hh)+1] # day plus 1
```

Number of tweets: `r nrow(sessionTweets)`

*Positive words...*

```{r}
#| warning: true

sessionTweets <- uniq_savingSessionTweetsDT[ba_created_at_date >= as.Date(start_hh) & # on the day
                                 ba_created_at_date <= as.Date(end_hh)+1] # day plus 1

# renders plot immediately (not sure why)
res <- makeSentimentCloud(sessionTweets, 
                   sentiFilter = "positive")
message("\nTweets from ", res$dates)
```

*Negative words...*

```{r}
#| warning: true

res <- makeSentimentCloud(sessionTweets, 
                   sentiFilter = "negative")
```

Total word and unique word counts...

```{r}
knitr::kable(res$ft, 
             caption = paste0("Number of unique words by sentiment")
             ) %>%
  kable_styling()

knitr::kable(res$nt, 
             caption = paste0("Number of words by sentiment")
             ) %>%
  kable_styling()

```

## Session 7 (NG-ESO 'live': `r pdt(savingSessions_dateTimes_DT[ID == 7, start_hh])`)

```{r}
start_hh <- savingSessions_dateTimes_DT[ID == 7, start_hh]
end_hh <- savingSessions_dateTimes_DT[ID == 7, end_hh]
actualEnd <- savingSessions_dateTimes_DT[ID == 7, end_hh] + 30*60
```

 * `r start_hh` to
 * `r actualEnd`

This was a more [generalised call](https://www.bbc.co.uk/news/business-64367504) for *DFS*

The same caveats about the comparisons apply here...

```{r}
#| label: makePlotSession7
#| tbl-cap: "MW comparisons"
#| #| warning: false

# catch if data not yet available
if(as.Date(start_hh) < Sys.Date()){
  res <- make_comparisonPlot(ngeso_dt_orig, start_hh, end_hh)
  print(res$p) # the plot
  
  makeFlexTable(res$wt[, hms:= as.factor(hms)], 
                cap = paste0("MW comparisons (hms = half hour period start, number of comparison days = ", res$nCompDays,")"),
                digits = 1) # the table
} else {
  message("Data not yet available, try later :-(")
}

```

### Carbon intensity

And just to keep in mind the carbon context:

```{r}
#| label: fig-seventhSavingSessionsCI
#| fig-cap: "Comparison of carbon intensity"
#| warning: false
if(as.Date(start_hh) < Sys.Date()){
res <- make_comparisonPlot(ngeso_dt_orig, start_hh, end_hh, 
                           yVar = "CARBON_INTENSITY",
                           yVarLab = "g CO2/kWh")
res$p
} else {
  message("Data not yet available, try later :-(")
}
```

Turning to hashtags, we have both #savingSessions and #demandFlexibilityService tweets for this event.

### #savingSessions tweets


```{r}

sessionTweets <- uniq_savingSessionTweetsDT[ba_created_at_date >= as.Date(start_hh) & # on the day
                                 ba_created_at_date <= as.Date(end_hh)+1] # day plus 1
```

Number of tweets: `r nrow(sessionTweets)`

*Positive words...*

```{r}
#| warning: true

# renders plot immediately (not sure why)
res <- makeSentimentCloud(sessionTweets, 
                   sentiFilter = "positive")
message("\nTweets from ", res$dates)
```

*Negative words...*

```{r}
#| warning: true

res <- makeSentimentCloud(sessionTweets, 
                   sentiFilter = "negative")
```

Total word and unique word counts...

```{r}
knitr::kable(res$ft, 
             caption = paste0("Number of unique words by sentiment")
             ) %>%
  kable_styling()

knitr::kable(res$nt, 
             caption = paste0("Number of words by sentiment")
             ) %>%
  kable_styling()

```


### #demandFlexibilityService tweets

```{r}
sessionTweets <- uniq_demandFlexibilityServiceTweetsDT[ba_created_at_date >= as.Date(start_hh) & # on the day
                                 ba_created_at_date <= as.Date(end_hh)+1] # day plus 1
```

Number of tweets: `r nrow(sessionTweets)`

*Positive words...*

```{r}
#| warning: true

# renders plot immediately (not sure why)
res <- makeSentimentCloud(sessionTweets, 
                   sentiFilter = "positive")
message("\nTweets from ", res$dates)
```

*Negative words...*

```{r}
#| warning: true

res <- makeSentimentCloud(sessionTweets, 
                   sentiFilter = "negative")
```

Total word and unique word counts...

```{r}
knitr::kable(res$ft, 
             caption = paste0("Number of unique words by sentiment")
             ) %>%
  kable_styling()

knitr::kable(res$nt, 
             caption = paste0("Number of words by sentiment")
             ) %>%
  kable_styling()

```


## Session 8 (NG-ESO 'live': `r pdt(savingSessions_dateTimes_DT[ID == 8, start_hh])`)

```{r}
start_hh <- savingSessions_dateTimes_DT[ID == 8, start_hh]
end_hh <- savingSessions_dateTimes_DT[ID == 8, end_hh]
actualEnd <- savingSessions_dateTimes_DT[ID == 8, end_hh] + 30*60
```

 * `r start_hh` to
 * `r actualEnd`

This was another [generalised call](https://www.bbc.co.uk/news/business-64372264) for *DFS*:

The same caveats about the comparisons apply here...

Also this event followed very soon after the last one so this will affect the comparison 'baseline' (previous n days).

```{r}
#| label: makePlotSession8
#| tbl-cap: "MW comparisons"

# catch if data not yet available
if(as.Date(start_hh) < Sys.Date()){
  res <- make_comparisonPlot(ngeso_dt_orig, start_hh, end_hh)
  print(res$p) # the plot
  
  makeFlexTable(res$wt[, hms:= as.factor(hms)], 
                cap = paste0("MW comparisons (hms = half hour period start, number of comparison days = ", res$nCompDays,")"),
                digits = 1) # the table
} else {
  message("Data not yet available, try later :-(")
}

```

### Carbon intensity

And just to keep in mind the carbon context:

```{r}
#| label: fig-eighthSavingSessionsCI
#| fig-cap: "Comparison of carbon intensity"

if(as.Date(start_hh) < Sys.Date()){
res <- make_comparisonPlot(ngeso_dt_orig, start_hh, end_hh, 
                           yVar = "CARBON_INTENSITY",
                           yVarLab = "g CO2/kWh")
res$p
} else {
  message("Data not yet available, try later :-(")
}
```

Turning to hashtags, we have both #savingSessions and #demandFlexibilityService tweets for this event.

### #savingSessions tweets

```{r}
sessionTweets <- uniq_savingSessionTweetsDT[ba_created_at_date >= as.Date(start_hh) & # on the day
                                 ba_created_at_date <= as.Date(start_hh)+1] # day plus 1
```

Number of tweets: `r nrow(sessionTweets)`

*Positive words...*

```{r}
#| warning: true

# renders plot immediately (not sure why)
res <- makeSentimentCloud(sessionTweets, 
                   sentiFilter = "positive")
message("\nTweets from ", res$dates)
```

*Negative words...*

```{r}
#| warning: true

res <- makeSentimentCloud(sessionTweets, 
                   sentiFilter = "negative")
```

Total word and unique word counts...

```{r}
knitr::kable(res$ft, 
             caption = paste0("Number of unique words by sentiment")
             ) %>%
  kable_styling()

knitr::kable(res$nt, 
             caption = paste0("Number of words by sentiment")
             ) %>%
  kable_styling()

```


### #demandFlexibilityService tweets

```{r}
sessionTweets <- uniq_demandFlexibilityServiceTweetsDT[ba_created_at_date >= as.Date(start_hh) & # on the day
                                 ba_created_at_date <= as.Date(end_hh)+1] # day plus 1
```

Number of tweets: `r nrow(sessionTweets)`

*Positive words...*

```{r}
#| warning: true

# renders plot immediately (not sure why)
res <- makeSentimentCloud(sessionTweets, 
                   sentiFilter = "positive")
message("\nTweets from ", res$dates)
```

*Negative words...*

```{r}
#| warning: true

res <- makeSentimentCloud(sessionTweets, 
                   sentiFilter = "negative")
```

Total word and unique word counts...

```{r}
knitr::kable(res$ft, 
             caption = paste0("Number of unique words by sentiment")
             ) %>%
  kable_styling()

knitr::kable(res$nt, 
             caption = paste0("Number of words by sentiment")
             ) %>%
  kable_styling()

```



## Session 9 (Octopus: `r pdt(savingSessions_dateTimes_DT[ID == 9, start_hh])`)

```{r}
start_hh <- savingSessions_dateTimes_DT[ID == 9, start_hh]
end_hh <- savingSessions_dateTimes_DT[ID == 9, end_hh]
actualEnd <- savingSessions_dateTimes_DT[ID == 9, end_hh] + 30*60
```

 * `r start_hh` to
 * `r actualEnd`

The same caveats about the comparisons apply here...

```{r}
#| label: makePlotSession9
#| tbl-cap: "MW comparisons"

# catch if data not yet available
if(as.Date(start_hh) < Sys.Date()){
  res <- make_comparisonPlot(ngeso_dt_orig, start_hh, end_hh)
  print(res$p) # the plot
  
  makeFlexTable(res$wt[, hms:= as.factor(hms)], 
                cap = paste0("MW comparisons (hms = half hour period start, number of comparison days = ", res$nCompDays,")"),
                digits = 1) # the table
} else {
  message("Data not yet available, try later :-(")
}

```

### Carbon intensity

And just to keep in mind the carbon context:

```{r}
#| label: fig-ninthSavingSessionsCI
#| fig-cap: "Comparison of carbon intensity"

if(as.Date(start_hh) < Sys.Date()){
res <- make_comparisonPlot(ngeso_dt_orig, start_hh, end_hh, 
                           yVar = "CARBON_INTENSITY",
                           yVarLab = "g CO2/kWh")
res$p
} else {
  message("Data not yet available, try later :-(")
}
```

Turning to hashtags, we have both #savingSessions and #demandFlexibilityService tweets for this event.

### #savingSessions tweets

```{r}
sessionTweets <- uniq_savingSessionTweetsDT[ba_created_at_date >= as.Date(start_hh) & # on the day
                                 ba_created_at_date <= as.Date(end_hh)+1] # day plus 1
```

Number of tweets: `r nrow(sessionTweets)`

*Positive words...*

```{r}
#| warning: true
# renders plot immediately (not sure why)
res <- makeSentimentCloud(sessionTweets, 
                   sentiFilter = "positive")
message("\nTweets from ", res$dates)
```

*Negative words...*

```{r}
#| warning: true

res <- makeSentimentCloud(sessionTweets, 
                   sentiFilter = "negative")
```

Total word and unique word counts...

```{r}
knitr::kable(res$ft, 
             caption = paste0("Number of unique words by sentiment")
             ) %>%
  kable_styling()

knitr::kable(res$nt, 
             caption = paste0("Number of words by sentiment")
             ) %>%
  kable_styling()

```


### #demandFlexibilityService tweets

Unlikely to be many of these as this was not a system-wide NG-ESO 'live' event.

```{r}
sessionTweets <- uniq_demandFlexibilityServiceTweetsDT[ba_created_at_date >= as.Date(start_hh) & # on the day
                                 ba_created_at_date <= as.Date(start_hh)+1] # day plus 1
```

Number of tweets: `r nrow(sessionTweets)`

*Positive words...*

```{r}
#| warning: true

# renders plot immediately (not sure why)
res <- makeSentimentCloud(sessionTweets, 
                   sentiFilter = "positive")
message("\nTweets from ", res$dates)
```

*Negative words...*

```{r}
#| warning: true

res <- makeSentimentCloud(sessionTweets, 
                   sentiFilter = "negative")
```

Total word and unique word counts...

```{r}
knitr::kable(res$ft, 
             caption = paste0("Number of unique words by sentiment")
             ) %>%
  kable_styling()

knitr::kable(res$nt, 
             caption = paste0("Number of words by sentiment")
             ) %>%
  kable_styling()

```




## Session 10 (Octopus: `r pdt(savingSessions_dateTimes_DT[ID == 10, start_hh])`)

```{r}
start_hh <- savingSessions_dateTimes_DT[ID == 10, start_hh]
end_hh <- savingSessions_dateTimes_DT[ID == 10, end_hh]
actualEnd <- savingSessions_dateTimes_DT[ID == 10, end_hh] + 30*60
```

 * `r start_hh` to
 * `r actualEnd`

The same caveats about the comparisons apply here...

```{r}
#| label: session_10
#| tbl-cap: "MW comparisons"
#| fig-cap: "MW comparisons"
#| 
# catch if data not yet available
if(as.Date(start_hh) < Sys.Date()){
  res <- make_comparisonPlot(ngeso_dt_orig, start_hh, end_hh)
  print(res$p) # the plot
  
  makeFlexTable(res$wt[, hms:= as.factor(hms)], 
                cap = paste0("MW comparisons (hms = half hour period start, number of comparison days = ", res$nCompDays,")"),
                digits = 1) # the table
} else {
  message("Data not yet available, try later :-(")
}

```

### Carbon intensity

And just to keep in mind the carbon context:

```{r}
#| label: fig-tenthSavingSessionsCI
#| fig-cap: "Comparison of carbon intensity"

if(as.Date(start_hh) < Sys.Date()){
res <- make_comparisonPlot(ngeso_dt_orig, start_hh, end_hh, 
                           yVar = "CARBON_INTENSITY",
                           yVarLab = "g CO2/kWh")
res$p
} else {
  message("Data not yet available, try later :-(")
}
```

Turning to hashtags, we have both #savingSessions and #demandFlexibilityService tweets for this event.

### #savingSessions tweets

```{r}
sessionTweets <- uniq_savingSessionTweetsDT[ba_created_at_date >= as.Date(start_hh) & # on the day
                                 ba_created_at_date <= as.Date(end_hh)+1] # day plus 1
```

Number of tweets: `r nrow(sessionTweets)`

*Positive words...*

```{r}
#| warning: true
# renders plot immediately (not sure why)
res <- makeSentimentCloud(sessionTweets, 
                   sentiFilter = "positive")
message("\nTweets from ", res$dates)
```

*Negative words...*

```{r}
#| warning: true

res <- makeSentimentCloud(sessionTweets, 
                   sentiFilter = "negative")
```

Total word and unique word counts...

```{r}
knitr::kable(res$ft, 
             caption = paste0("Number of unique words by sentiment")
             ) %>%
  kable_styling()

knitr::kable(res$nt, 
             caption = paste0("Number of words by sentiment")
             ) %>%
  kable_styling()

```


### #demandFlexibilityService tweets

Unlikely to be many of these as this was not a system-wide NG-ESO 'live' event.

```{r}
sessionTweets <- uniq_demandFlexibilityServiceTweetsDT[ba_created_at_date >= as.Date(start_hh) & # on the day
                                 ba_created_at_date <= as.Date(start_hh)+1] # day plus 1
```

Number of tweets: `r nrow(sessionTweets)`

*Positive words...*

```{r}
#| warning: true
# breaks if no data yet
# renders plot immediately (not sure why)
try(res <- makeSentimentCloud(sessionTweets, 
                   sentiFilter = "positive"))
message("\nTweets from ", res$dates)
```

*Negative words...*

```{r}
#| warning: true
# breaks if no data yet
try(res <- makeSentimentCloud(sessionTweets, 
                   sentiFilter = "negative"))
```

Total word and unique word counts...

```{r}
knitr::kable(res$ft, 
             caption = paste0("Number of unique words by sentiment")
             ) %>%
  kable_styling()

knitr::kable(res$nt, 
             caption = paste0("Number of words by sentiment")
             ) %>%
  kable_styling()

```



## Session 11 (Octopus: `r pdt(savingSessions_dateTimes_DT[ID == 11, start_hh])`)

```{r}
start_hh <- savingSessions_dateTimes_DT[ID == 11, start_hh]
end_hh <- savingSessions_dateTimes_DT[ID == 11, end_hh]
actualEnd <- savingSessions_dateTimes_DT[ID == 11, end_hh] + 30*60
```

 * `r start_hh` to
 * `r actualEnd`

The same caveats about the comparisons apply here...

```{r}
#| label: session_11
#| tbl-cap: "MW comparisons"
#| fig-cap: "MW comparisons"
#| 
# catch if data not yet available
if(as.Date(start_hh) < Sys.Date()){
  res <- make_comparisonPlot(ngeso_dt_orig, start_hh, end_hh)
  print(res$p) # the plot
  
  makeFlexTable(res$wt[, hms:= as.factor(hms)], 
                cap = paste0("MW comparisons (hms = half hour period start, number of comparison days = ", res$nCompDays,")"),
                digits = 1) # the table
} else {
  message("Data not yet available, try later :-(")
}

```

### Carbon intensity

And just to keep in mind the carbon context:

```{r}
#| label: fig-eleventhSavingSessionsCI
#| fig-cap: "Comparison of carbon intensity"

if(as.Date(start_hh) < Sys.Date()){
res <- make_comparisonPlot(ngeso_dt_orig, start_hh, end_hh, 
                           yVar = "CARBON_INTENSITY",
                           yVarLab = "g CO2/kWh")
res$p
} else {
  message("Data not yet available, try later :-(")
}
```

Turning to hashtags, we have both #savingSessions and #demandFlexibilityService tweets for this event.

### #savingSessions tweets

```{r}
sessionTweets <- uniq_savingSessionTweetsDT[ba_created_at_date >= as.Date(start_hh) & # on the day
                                 ba_created_at_date <= as.Date(end_hh)+1] # day plus 1
```

Number of tweets: `r nrow(sessionTweets)`

*Positive words...*

```{r}
#| warning: true
# renders plot immediately (not sure why)
res <- makeSentimentCloud(sessionTweets, 
                   sentiFilter = "positive")
message("\nTweets from ", res$dates)
```

*Negative words...*

```{r}
#| warning: true

res <- makeSentimentCloud(sessionTweets, 
                   sentiFilter = "negative")
```

Total word and unique word counts...

```{r}
knitr::kable(res$ft, 
             caption = paste0("Number of unique words by sentiment")
             ) %>%
  kable_styling()

knitr::kable(res$nt, 
             caption = paste0("Number of words by sentiment")
             ) %>%
  kable_styling()

```


### #demandFlexibilityService tweets

Unlikely to be many of these as this was not a system-wide NG-ESO 'live' event.

```{r}
sessionTweets <- uniq_demandFlexibilityServiceTweetsDT[ba_created_at_date >= as.Date(start_hh) & # on the day
                                 ba_created_at_date <= as.Date(start_hh)+1] # day plus 1
```

Number of tweets: `r nrow(sessionTweets)`

*Positive words...*

```{r}
#| warning: true
# breaks if no data yet
# renders plot immediately (not sure why)
try(res <- makeSentimentCloud(sessionTweets, 
                   sentiFilter = "positive"))
message("\nTweets from ", res$dates)
```

*Negative words...*

```{r}
#| warning: true
# breaks if no data yet
try(res <- makeSentimentCloud(sessionTweets, 
                   sentiFilter = "negative"))
```

Total word and unique word counts...

```{r}
knitr::kable(res$ft, 
             caption = paste0("Number of unique words by sentiment")
             ) %>%
  kable_styling()

knitr::kable(res$nt, 
             caption = paste0("Number of words by sentiment")
             ) %>%
  kable_styling()

```



## Session 12 (NG-ESO 'live' (?): `r pdt(savingSessions_dateTimes_DT[ID == 12, start_hh])`)

```{r}
start_hh <- savingSessions_dateTimes_DT[ID == 12, start_hh]
end_hh <- savingSessions_dateTimes_DT[ID == 12, end_hh]
actualEnd <- savingSessions_dateTimes_DT[ID == 12, end_hh] + 30*60
```

 * `r start_hh` to
 * `r actualEnd`

The same caveats about the comparisons apply here...

```{r}
#| tbl-cap: "MW comparisons"
#| fig-cap: "MW comparisons"
#| 
# catch if data not yet available
if(as.Date(start_hh) < Sys.Date()){
  res <- make_comparisonPlot(ngeso_dt_orig, start_hh, end_hh)
  print(res$p) # the plot
  
  makeFlexTable(res$wt[, hms:= as.factor(hms)], 
                cap = paste0("MW comparisons (hms = half hour period start, number of comparison days = ", res$nCompDays,")"),
                digits = 1) # the table
} else {
  message("Data not yet available, try later :-(")
}

```

### Carbon intensity

And just to keep in mind the carbon context:

```{r}
#| fig-cap: "Comparison of carbon intensity"

if(as.Date(start_hh) < Sys.Date()){
res <- make_comparisonPlot(ngeso_dt_orig, start_hh, end_hh, 
                           yVar = "CARBON_INTENSITY",
                           yVarLab = "g CO2/kWh")
res$p
} else {
  message("Data not yet available, try later :-(")
}
```

Turning to hashtags, we have both #savingSessions and #demandFlexibilityService tweets for this event.

### #savingSessions tweets

```{r}
sessionTweets <- uniq_savingSessionTweetsDT[ba_created_at_date >= as.Date(start_hh) & # on the day
                                 ba_created_at_date <= as.Date(end_hh)+1] # day plus 1
```

Number of tweets: `r nrow(sessionTweets)`

*Positive words...*

```{r}
#| warning: true
# renders plot immediately (not sure why)
res <- makeSentimentCloud(sessionTweets, 
                   sentiFilter = "positive")
message("\nTweets from ", res$dates)
```

*Negative words...*

```{r}
#| warning: true

res <- makeSentimentCloud(sessionTweets, 
                   sentiFilter = "negative")
```

Total word and unique word counts...

```{r}
knitr::kable(res$ft, 
             caption = paste0("Number of unique words by sentiment")
             ) %>%
  kable_styling()

knitr::kable(res$nt, 
             caption = paste0("Number of words by sentiment")
             ) %>%
  kable_styling()

```


### #demandFlexibilityService tweets


```{r}
sessionTweets <- uniq_demandFlexibilityServiceTweetsDT[ba_created_at_date >= as.Date(start_hh) & # on the day
                                 ba_created_at_date <= as.Date(start_hh)+1] # day plus 1
```

Number of tweets: `r nrow(sessionTweets)`

*Positive words...*

```{r}
#| warning: true
# breaks if no data yet
# renders plot immediately (not sure why)
try(res <- makeSentimentCloud(sessionTweets, 
                   sentiFilter = "positive"))
message("\nTweets from ", res$dates)
```

*Negative words...*

```{r}
#| warning: true
# breaks if no data yet
try(res <- makeSentimentCloud(sessionTweets, 
                   sentiFilter = "negative"))
```

Total word and unique word counts...

```{r}
knitr::kable(res$ft, 
             caption = paste0("Number of unique words by sentiment")
             ) %>%
  kable_styling()

knitr::kable(res$nt, 
             caption = paste0("Number of words by sentiment")
             ) %>%
  kable_styling()

```


# Future work

 * more trial days
 * use 5 days before/after as comparators? Confounded by events close together...

```{r}
#| label: theEnd
```
