---
title: "savingSessions: energy analysis"
author: "Ben Anderson (@dataknut)"
date: 'Last run at: `r Sys.time()`'
format: 
  html:
    self-contained: true
    toc: true
    toc-depth: 3
    code-fold: true
    number-sections: true
execute:
  echo: false
  warning: false
editor: visual
---

# Background

UK demand response experiments by NG-ESO and retailers such as [\@OctopusEnergy](https://twitter.com/SavingSessions)

How big a 'dent' in demand did they make?

# Code setup

Part of <https://github.com/dataknut/savingSessions>

```{r}
#| label: codeSetup
#| warning: false

library(data.table)
library(dkUtils)
library(ggplot2)
library(hms)
library(lubridate)

```

# Getting data

Grab the most NG-ESO 'gen mix' [data](https://data.nationalgrideso.com/carbon-intensity1/historic-generation-mix).

> Should probably use the [demand data](https://data.nationalgrideso.com/demand/historic-demand-data) instead as it includes embedded wind & solar.

```{r}
#| label: getData
#| warning: false

esoF <- here::here("data", "latest_df_fuel_ckan.csv")
if(file.exists(paste0(esoF, ".gz"))){
  message("We already have a version saved to: ", paste0(esoF, ".gz"))
  message("Loading it...")
  ngeso_dt_orig <- data.table::fread(paste0(esoF, ".gz"))
} else {
  message("We don't already have a version, downloading and saving to: ", esoF)
  ngeso_dt_orig <- data.table::fread("https://data.nationalgrideso.com/backend/dataset/88313ae5-94e4-4ddc-a790-593554d8c6b9/resource/f93d1835-75bc-43e5-84ad-12472b180a98/download/df_fuel_ckan.csv")
  # nice dateTime
  ngeso_dt_orig[, dv_start := lubridate::as_datetime(DATETIME)]
  data.table::fwrite(ngeso_dt_orig, esoF)
  dkUtils::gzipIt(esoF)
  data.table::fwrite(ngeso_dt_orig, "~/Dropbox/data/UK_NGESO/genMix/latest_df_fuel_ckan.csv") # save locally for future re-use
  dkUtils::gzipIt("~/Dropbox/data/UK_NGESO/genMix/latest_df_fuel_ckan.csv")
}

# if older than 1 day, reload
today <- lubridate::today()
lastNGESO <- as.Date(max(ngeso_dt_orig$dv_start))

if(today - lastNGESO > 1) {
  # old data, reload
  message("But the version we have dates from ", lastNGESO, " (",today - lastNGESO ," days ago), downloading latest...")
  ngeso_dt_orig <- data.table::fread("https://data.nationalgrideso.com/backend/dataset/88313ae5-94e4-4ddc-a790-593554d8c6b9/resource/f93d1835-75bc-43e5-84ad-12472b180a98/download/df_fuel_ckan.csv")
  # nice dateTime
  ngeso_dt_orig[, dv_start := lubridate::as_datetime(DATETIME)]
  data.table::fwrite(ngeso_dt_orig, esoF)
  dkUtils::gzipIt(esoF)
}

# we think renewable is wind + solar, low carbon includes nuclear
```

Select the dates and periods we need. GENERATION is in MW. It has some exclusions, its just an indicator.

```{r}
#|label: selectSessionDates

session1DateStart <- lubridate::as_datetime("2022-11-15 17:00:00") # the half-hour it starts
session1DateEnd <- lubridate::as_datetime("2022-11-15 18:00:00") # 

session2DateStart <- lubridate::as_datetime("2022-11-22 17:30:00") # the half-hour it starts
session2DateEnd <- lubridate::as_datetime("2022-11-22 18:30:00") # 
```

1.  Session 1: `r session1DateStart` to `r session1DateEnd`

2.  Session 2: `r session2DateStart` to `r session2DateEnd`

```{r}
session1_dt <- ngeso_dt_orig[dv_start >= session1DateStart & 
                               dv_start < session1DateEnd]
session2_dt <- ngeso_dt_orig[dv_start >= session2DateStart & 
                               dv_start < session2DateEnd]

message("Session 1:")
head(session1_dt[, .(dv_start, GENERATION)])

message("Session 2:")
head(session2_dt[, .(dv_start, GENERATION)])

ngeso_dt_orig[, dv_date := lubridate::date(dv_start)]
sessions_dt <- ngeso_dt_orig[dv_date == lubridate::date(session1DateStart) |
                               dv_date == lubridate::date(session2DateStart)]

sessions_dt[, hms := hms::as_hms(dv_start)]

periodAlpha <- 0.3 #  shaded rects on plots
periodFill <- "grey50"
plotDT <- sessions_dt[dv_date == lubridate::date(session1DateStart)]
                      
ymax <- max(plotDT$GENERATION/1000)
ymin <- 0

psess1 <- ggplot2::ggplot(plotDT, 
                          aes(x = hms, y = GENERATION/1000)) +
  geom_line() +
  ylim(0, NA) +
  facet_grid(. ~ dv_date) +
    annotate("rect", xmin = hms::as_hms(session1DateStart),
           xmax = hms::as_hms(session1DateEnd),
           ymin = ymin, ymax = ymax,
           alpha = periodAlpha, fill = periodFill) +
  labs(x = "Time of day",
       y = "GENERATION (GW)",
       caption = "Session 1: NG-ESO Generation Mix data (saving session marked)")
psess1

plotDT <- sessions_dt[dv_date == lubridate::date(session2DateStart)]
                      
ymax <- max(plotDT$GENERATION/1000)
ymin <- 0
psess2 <- ggplot2::ggplot(plotDT, 
                          aes(x = hms, y = GENERATION/1000)) +
  geom_line() +
  ylim(0, NA) +
  facet_grid(. ~ dv_date) +
    annotate("rect", xmin = hms::as_hms(session2DateStart),
           xmax = hms::as_hms(session2DateEnd),
           ymin = ymin, ymax = ymax,
           alpha = periodAlpha, fill = periodFill) +
   labs(x = "Time of day",
       y = "GENERATION (GW)",
       caption = "Session 2: NG-ESO Generation Mix data (saving session marked)")
psess2
```

# Analysis

## Session 1

According to <https://energycentral.com/news/octopus-energy-provide-108mw-grid-flexibility-first-%E2%80%98saving-session%E2%80%99>:

>"Over 200,000 households reduced their energy demand by 108MW collectively, the same as a gas power station can generate in an hour. If replicated by all UK energy suppliers, this would be over 1GW."

But <https://www.current-news.co.uk/news/octopus-energy-customers-provide-108mw-of-flexibility-during-first-saving-session> quotes Octopus as actually saying: 

>"if the program was scaled to all electric smart meter customers with all supplier[s] in Great Britain, it could create over 1GW of flexible energy load"


```{r}
session1_mean_mw <- mean(session1_dt$GENERATION)
session1_mean_gw <- session1_mean_mw/1000

scaled_kW <- (108/200)*28000000
scaled_GW <- ((108/200)*28000000)/1000000

scaled_GW_pc <- 100*(scaled_GW/session1_mean_gw)

gb_smartMeters <- 14000000
mean_kW_per_participant <- 108/200
scaled_GB_smart_meters_GW_pc <- 100*((gb_smartMeters*mean_kW_per_participant)/1000000)/session1_mean_gw
```

Octopus customers' 108 MW was `r round(100*(108/session1_mean_mw),3)` % of mean GB generation (`r session1_mean_gw` GW) over the hour of the session (see NG-ESO data above).

108 MW over 200,000 customers is `r 108/200000` MW per household or, more sensibly, `r mean_kW_per_participant` kW per household. A pretty reasonable reduction.

But if 200,000 produced a 108 MW power reduction (`r mean_kW_per_participant` kW per household ), the 1 GW estimate implies we only have `r prettyNum(1000000/mean_kW_per_participant, big.mark = ",")` GB households with a smart meter. Really? Or did they mean all suppliers taking part in the scheme?

Anyway, the latest data has 14 million domestic eletricity smart meters as of [Q1 2022](https://assets.publishing.service.gov.uk/government/uploads/system/uploads/attachment_data/file/1099629/Q2_2022_Smart_Meters_Statistics_Report.pdf) - 45% of all domestic meters.


So:

  * If all [~28 million households](https://www.ons.gov.uk/peoplepopulationandcommunity/birthsdeathsandmarriages/families/bulletins/familiesandhouseholds/2021) in the UK did this it would be `r scaled_kW` kW. That's `r scaled_GW` GW...
  * If we reduce ambition a bit and assume that only the 14 million domestic GB customers with a smart meter as of Q2 2022 did this then that would provide `r (gb_smartMeters*mean_kW_per_participant)/1000000` GW. Which would be `r round(scaled_GB_smart_meters_GW_pc,2)` % of generation. Not bad aye?

The trouble is... these are octopus customers self-selecting into a trial. You just have to [read the tweets](https://twitter.com/SavingSessions) to see that there are households with installed batteries, pausable heat pumps in well-insulated homes and EV V2G. At what point will we all be similarly equipped?

## Session 2

# The end

```{r}
#| label: theEnd
```
