---
title: "savingSessions: tweet analysis"
author: "Ben Anderson (@dataknut)"
date: 'Last run at: `r Sys.time()`'
format: 
  html:
    self-contained: true
    toc: true
    toc-depth: 3
    code-fold: true
    number-sections: true
execute:
  echo: false
  warning: false
editor: visual
---

# Background

UK demand response experiments by NG-ESO and retailers such as [\@OctopusEnergy](https://twitter.com/SavingSessions)

Attempt to do some analysis of #savingSession(s) tweets.

# Code setup

Part of <https://github.com/dataknut/savingSessions>

Makes use of <https://github.com/dataknut/hashTagR>, a DIY wrapper for the [rtweet](https://docs.ropensci.org/rtweet) rstats package.

```{r}
#| label: codeSetup
#| warning: false

library(data.table)
library(hashTagR)
library(ggplot2)
library(lubridate)
library(readr)
library(rtweet)

hashtags <- c("savingSession", "savingSessions", "savingsession") # fairly sure it's case insensitive
searchString <- hashTagR::createSearchFromTags(hashtags) # convert to rtweet search string

#searchString <- "#SavingSession OR #SavingSessions" # edit to suit

dataPath <- "~/Dropbox/data/twitter/savingSessions/" # edit to suit

session1DateStart <- lubridate::as_datetime("2022-11-15 17:00:00") # the half-hour it starts
session1DateEnd <- lubridate::as_datetime("2022-11-15 18:00:00") # 

session2DateStart <- lubridate::as_datetime("2022-11-22 17:30:00") # the half-hour it starts
session2DateEnd <- lubridate::as_datetime("2022-11-22 18:30:00") # 


session3DateStart <- lubridate::as_datetime("2022-11-30 17:30:00") # the half-hour it starts
session3DateEnd <- lubridate::as_datetime("2022-11-30 18:30:00") # 


session4DateStart <- lubridate::as_datetime("2022-12-01 17:00:00") # the half-hour it starts
session4DateEnd <- lubridate::as_datetime("2022-12-01 18:00:00") # 

# functions ----

add_sessionDate <- function(p){
  # add session dates to plot
  p <- p +
  annotate("rect", xmin = session1DateStart,
           xmax = session1DateEnd,
           ymin = ymin, ymax = ymax,
           alpha = periodAlpha, fill = periodFill) +
    annotate("text", x = session1DateStart-(24*60*60), 
             y = ymax*0.9, label = paste0("Session 1\n",
                                          lubridate::as_date(session1DateStart))) + 
  annotate("rect", xmin = session2DateStart,
           xmax = session2DateEnd,
           ymin = ymin, ymax = ymax,
           alpha = periodAlpha, fill = periodFill) +
        annotate("text", x = session2DateStart-(24*60*60), 
             y = ymax*0.9, label = paste0("Session 2\n",
                                          lubridate::as_date(session2DateStart))) + 
  annotate("rect", xmin = session3DateStart,
           xmax = session3DateEnd,
           ymin = ymin, ymax = ymax,
           alpha = periodAlpha, fill = periodFill) +
        annotate("text", x = session3DateStart-(24*60*60), 
             y = ymax*0.9, label = paste0("Session 3\n",
                                          lubridate::as_date(session3DateStart)))  +
    annotate("rect", xmin = session4DateStart,
           xmax = session4DateEnd,
           ymin = ymin, ymax = ymax,
           alpha = periodAlpha, fill = periodFill) +
        annotate("text", x = session4DateStart-(24*60*60), 
             y = ymax*0.8, label = paste0("Session 4\n",
                                          lubridate::as_date(session4DateStart)))
  return(p)
}
```

# Getting data

Grab the most recent set of tweets that mention `r searchString` using the [rtweet::search_tweet()](https://docs.ropensci.org/rtweet/reference/search_tweets.html) function and merge with any we may already have downloaded.

Note that tweets do not seem to be available after \~ 14 days via the API used by rtweet. Best to keep refreshing the data every week...

```{r}
#| label: getData
#| warning: false

now <- lubridate::now()

if(dir.exists(dataPath)){
  ofile <- path.expand(paste0(dataPath, "tw_",now,".csv"))
  tweetsDT <- hashTagR::getTweets(searchString, n = 20000)
  readr::write_csv(tweetsDT, file = ofile) # data.table breaks here
  message("Retrieved ", nrow(tweetsDT), " tweets and saved them to ", ofile)
} else {
  message(dataPath, " not found!")
}

# now load all the tweet files we have matching that searchString

dt <- hashTagR::loadTweets(path = dataPath, pattern = "*.csv")

u_dt <- hashTagR::processTweets(dt)
```

That produced a data file of `r nrow(u_dt)` tweets.

We do NOT store the tweets in the repo for both [ethical](https://blogs.lse.ac.uk/impactofsocialsciences/2015/09/28/challenges-of-using-twitter-as-a-data-source-resources/) and practical reasons...

Note also that we may not be collecting the complete dataset of hashtagged tweets due to the [intricacies of the twitter API](https://www.demos.co.uk/files/Road_to_representivity_final.pdf?1441811336).

# Analysis

## Tweet time line

@fig-timeLine shows the timing of tweets by hour.

```{r}
#| echo: false
#| label: fig-timeLine
#| fig-cap: "Tweets over time"

plotDT <- u_dt[, .(nObs = .N), keyby = .(Time = ba_created_at_dateHour)]

periodAlpha <- 0.3 #  shaded rects on plots
periodFill <- "grey50"
ymax <- max(plotDT$nObs)
ymin <- min(plotDT$nObs)


# plot over time

myCaption <- "Data: twitter API via @mkearney's rtweet package\nPlot by @dataknut\nSaving Session period(s) marked as grey box"
p <- ggplot2::ggplot(plotDT, aes(x = Time, y = nObs)) + 
  geom_point() +
  labs(caption = myCaption)
add_sessionDate(p)
```

@fig-cumulativeTimeLine shows cumulative tweets by hour.

```{r}
#| echo: false
#| label: fig-cumulativeTimeLine
#| fig-cap: "Cumulative number of tweets over time"


plotDT[, cumSum := cumsum(nObs)]
ymax <- max(plotDT$cumSum)
ymin <- min(plotDT$cumSum)
p <- ggplot2::ggplot(plotDT, aes(x = Time, y = cumSum)) + 
  geom_point() +
  geom_line() +
  labs(caption = myCaption)

add_sessionDate(p)
```

We see roughly the kind of uptick in tweets for Session 2 that we saw for Session 1...

## Sentiment analysis

Yeah...

## Content analysis

Yeah...

# The end

```{r}
#| label: theEnd
```
