---
title: "savingSessions: tweet analysis"
author: "Ben Anderson (@dataknut)"
date: 'Last run at: `r Sys.time()`'
format: 
  html:
    self-contained: true
    toc: true
    toc-depth: 3
    code-fold: true
    number-sections: true
execute:
  echo: false
  warning: false
editor: visual
---

# Background

UK demand response experiments by NG-ESO and retailers such as [\@OctopusEnergy](https://twitter.com/SavingSessions)

Attempt to do some analysis of #savingSession(s) tweets.

Inspired by <https://docs.ropensci.org/rtweet/>

Last run at: `r Sys.time()`

# Setup

::: panel-tabset
## Code setup

Part of <https://github.com/dataknut/savingSessions>

Makes use of <https://github.com/dataknut/hashTagR>, a DIY wrapper for the [rtweet](https://docs.ropensci.org/rtweet) rstats package.

```{r}
#| label: codeSetup
#| warning: false

myLibs <- c("data.table",
            "dplyr",
            "hashTagR",
            "ggplot2",
            "knitr",
            "kableExtra",
            "lubridate",
            "readr",
            "rtweet",
            "tidytext",
            "wordcloud")

# load the libraries, install if we don't have them
dkUtils::loadLibraries(myLibs) # get it here: devtools::install_github("dataknut/dkUtils")

hashtags <- c("savingSession", "savingSessions", "savingsession") # fairly sure it's case insensitive
searchString <- hashTagR::createSearchFromTags(hashtags) # convert to rtweet search string

#searchString <- "#SavingSession OR #SavingSessions" # edit to suit

# ovo call it Shift & Save https://twitter.com/OVOEnergy/status/1597591031507476486

dataPath <- "~/Dropbox/data/twitter/savingSessions/" # edit to suit

session1DateStart <- lubridate::as_datetime("2022-11-15 17:00:00") # the half-hour it starts
session1DateEnd <- lubridate::as_datetime("2022-11-15 18:00:00") # 

session2DateStart <- lubridate::as_datetime("2022-11-22 17:30:00") # the half-hour it starts
session2DateEnd <- lubridate::as_datetime("2022-11-22 18:30:00") # 

session3DateStart <- lubridate::as_datetime("2022-11-30 17:30:00") # the half-hour it starts
session3DateEnd <- lubridate::as_datetime("2022-11-30 18:30:00") # 

session4DateStart <- lubridate::as_datetime("2022-12-01 17:00:00") # the half-hour it starts
session4DateEnd <- lubridate::as_datetime("2022-12-01 18:00:00") # 

session5DateStart <- lubridate::as_datetime("2022-12-12 17:00:00") # the half-hour it starts
session5DateEnd <- lubridate::as_datetime("2022-12-12 18:30:00") # 

session6DateStart <- lubridate::as_datetime("2023-01-19 09:00:00") # the half-hour it starts
session6DateEnd <- lubridate::as_datetime("2023-01-19 09:30:00") # 

session7DateStart <- lubridate::as_datetime("2023-01-23 17:00:00") # the half-hour it starts
session7DateEnd <- lubridate::as_datetime("2023-01-23 17:30:00") # 

# functions ----

add_sessionDates <- function(p){
  # add session dates to plot
  p <- p +
  annotate("rect", xmin = session1DateStart,
           xmax = session1DateEnd,
           ymin = ymin, ymax = ymax,
           alpha = periodAlpha, fill = periodFill) +
    annotate("text", x = session1DateStart-(24*60*60), vjust=-0.1, angle=90,
             y = ymax*0.5, label = paste0("Session 1: ", lubridate::as_date(session1DateStart))) + 
  annotate("rect", xmin = session2DateStart,
           xmax = session2DateEnd,
           ymin = ymin, ymax = ymax,
           alpha = periodAlpha, fill = periodFill) +
        annotate("text", x = session2DateStart-(24*60*60), vjust=-0.1, angle=90,
             y = ymax*0.5, label = paste0("Session 2: ",
                                          lubridate::as_date(session2DateStart))) + 
  annotate("rect", xmin = session3DateStart,
           xmax = session3DateEnd,
           ymin = ymin, ymax = ymax,
           alpha = periodAlpha, fill = periodFill) +
        annotate("text", x = session3DateStart-(24*60*60), vjust=-0.1, angle=90,
             y = ymax*0.5, label = paste0("Session 3: ",
                                          lubridate::as_date(session3DateStart)))  +
    annotate("rect", xmin = session4DateStart,
           xmax = session4DateEnd,
           ymin = ymin, ymax = ymax,
           alpha = periodAlpha, fill = periodFill) +
        annotate("text", x = session4DateStart+(48*60*60), vjust=-0.1, angle=90,
             y = ymax*0.5, label = paste0("Session 4: ",
                                          lubridate::as_date(session4DateStart))) +
      annotate("rect", xmin = session5DateStart,
           xmax = session5DateEnd,
           ymin = ymin, ymax = ymax,
           alpha = periodAlpha, fill = periodFill) +
        annotate("text", x = session5DateStart-(24*60*60), vjust=-0.1, angle=90,
             y = ymax*0.5, label = paste0("Session 5: ",
                                          lubridate::as_date(session5DateStart))) +
  annotate("rect", xmin = session6DateStart,
           xmax = session6DateEnd,
           ymin = ymin, ymax = ymax,
           alpha = periodAlpha, fill = periodFill) +
        annotate("text", x = session6DateStart-(24*60*60), vjust=-0.1, angle=90,
             y = ymax*0.5, label = paste0("Session 6: ",
                                          lubridate::as_date(session6DateStart)))
  return(p)
}
```

## Getting data

Grab the most recent set of tweets that mention `r searchString` using the [rtweet::search_tweet()](https://docs.ropensci.org/rtweet/reference/search_tweets.html) function and merge with any we may already have downloaded.

> Should we also try to get all replies to @savingSessions?

Note that tweets do not seem to be available after \~ 14 days via the API used by rtweet. Best to keep refreshing the data every week...

```{r}
#| label: getData

now <- lubridate::now()

if(dir.exists(dataPath)){
  ofile <- path.expand(paste0(dataPath, "tw_",now,".csv"))
  tweetsDT <- hashTagR::getTweets(searchString, n = 20000)
  readr::write_csv(tweetsDT, file = ofile) # data.table breaks here
  message("Retrieved ", nrow(tweetsDT), " tweets and saved them to ", ofile)
} else {
  message(dataPath, " not found!")
}

# now load all the tweet files we have matching that searchString

tweetsDT <- hashTagR::loadTweets(path = dataPath, pattern = "*.csv")

uniq_tweetsDT <- hashTagR::processTweets(tweetsDT)
```

That produced a data file of `r nrow(uniq_tweetsDT)` tweets.

We do NOT store the tweets in the repo for both [ethical](https://blogs.lse.ac.uk/impactofsocialsciences/2015/09/28/challenges-of-using-twitter-as-a-data-source-resources/) and practical reasons...

Note also that we may not be collecting the complete dataset of hashtagged tweets due to the [intricacies of the twitter API](https://www.demos.co.uk/files/Road_to_representivity_final.pdf?1441811336).
:::

# Analysis

::: panel-tabset

## Tweet time line

@fig-timeLine shows the timing of tweets by hour.

```{r}
#| echo: false
#| label: fig-timeLine
#| fig-cap: "Tweets over time"

plotDT <- uniq_tweetsDT[, .(nObs = .N), keyby = .(Time = ba_created_at_dateHour)]

periodAlpha <- 0.3 #  shaded rects on wc_plots
periodFill <- "red"
ymax <- max(plotDT$nObs)
ymin <- min(plotDT$nObs)


# plot over time

myCaption <- "Data: twitter API via @mkearney's rtweet package\nPlot by @dataknut\nSaving Session period(s) marked"
p <- ggplot2::ggplot(plotDT, aes(x = Time, y = nObs)) + 
  geom_point() +
  labs(caption = myCaption)
add_sessionDates(p)
```

@fig-cumulativeTimeLine shows cumulative tweets by hour.

```{r}
#| echo: false
#| label: fig-cumulativeTimeLine
#| fig-cap: "Cumulative number of tweets over time"


plotDT[, cumSum := cumsum(nObs)]
ymax <- max(plotDT$cumSum)
ymin <- min(plotDT$cumSum)
p <- ggplot2::ggplot(plotDT, aes(x = Time, y = cumSum)) + 
  geom_point() +
  geom_line() +
  labs(caption = myCaption)

add_sessionDates(p)
```

We see roughly the kind of uptick in tweets for Session 2 that we saw for Session 1...

## Content analysis

Let's try a word cloud.

Inspiration here: <https://towardsdatascience.com/create-a-word-cloud-with-r-bde3e7422e8a>

```{r}
#| label: cleanTweets

# clean up the tweet text to remove stuff we don't want.
uniq_tweetsDT[, clean_text := gsub("https\\S*", " ", text)]
uniq_tweetsDT[, clean_text := gsub("@\\S*", " ", clean_text)]
uniq_tweetsDT[, clean_text := gsub("amp", " ", clean_text)]
uniq_tweetsDT[, clean_text := gsub("RT", " ", clean_text)]
uniq_tweetsDT[, clean_text := gsub("[\r\n]", " ", clean_text)]
uniq_tweetsDT[, clean_text := gsub("[[:punct:]]", " ", clean_text)]

```

```{r}
#| label: makeWords

tweets_words <-  uniq_tweetsDT %>%
 select(clean_text) %>%
 unnest_tokens(word, clean_text)

words <- tweets_words %>% count(word, sort=TRUE)

tweets_words_per_day <-  uniq_tweetsDT %>%
  group_by(ba_created_at_date) %>%
 select(clean_text) %>%
 unnest_tokens(word, clean_text)

words_per_day <- tweets_words_per_day %>% count(word, sort=TRUE)
```

Make a word cloud for all words in all tweets.

These *may* not render the word 'savingsession' as it will be in all tweets due to the twitter search pattern used.

```{r}
#| label: makeCloud

set.seed(1234) # for reproducibility 
wordcloud::wordcloud(words = words$word, 
          freq = words$n, min.freq = 1,           
          max.words=200, 
          random.order=FALSE, 
          rot.per=0.35,            
          colors=brewer.pal(8, "Dark2"))
```

We need to remove common words (to, the, and, a, for, etc). These are called 'stop words'.

What happens if we do that?

```{r}
#| label: cleanWordCloud

data(stop_words)

# remove the stop words - they are usually the most frequent
# and usually the least interesting
reduced_df <- words %>%
  dplyr::anti_join(stop_words)

reduced_df_daily <- words_per_day %>%
  dplyr::anti_join(stop_words)

# redraw
set.seed(1234) # for reproducibility 
wordcloud::wordcloud(words = reduced_df$word, 
                     freq = reduced_df$n, min.freq = 1,           
                     max.words=200, 
                     random.order=FALSE, 
                     rot.per=0.35,            
                     colors=brewer.pal(8, "Dark2"))
```

Not especially informative... Perhaps we should try to extract the 'sentiment' of the words.

## Sentiment analysis (all tweets)

Inspired by <https://www.tidytextmining.com/sentiment.html>

Take those cleaned words and sentiment them!

In each case we show the number of negative and positive codings for the unique words (which will add up to the Number of unique words by sentiment) and then the total frequency of words that are negative or positive (which will add up to the total number of words).

> Got it?

The first word cloud shows names that have negative sentiment (according to `tidytext::get_sentiments("bing")`). Remember the size of the words is relative to the count of all negative words.

```{r}
#| label: sentiment
bing_sentiment_counts <- reduced_df %>%
  inner_join(tidytext::get_sentiments("bing")) %>% # only keeps those with sentiments
  #count(linenumber, sentiment, sort = TRUE) %>%
  ungroup()

bing_sentiment_counts_daily <- reduced_df_daily %>%
  inner_join(tidytext::get_sentiments("bing")) %>% # only keeps those with sentiments
  #count(linenumber, sentiment, sort = TRUE) %>%
  ungroup()

message("\nNumber of unique words by sentiment")
uniqueN(bing_sentiment_counts$word)

message("\nDo we have roughly the same number of each kind of word?")
table(bing_sentiment_counts$sentiment)

message("\nNumber of words coded")
sum(bing_sentiment_counts$n)

message("What about the overall frequency of these words?")
bing_sentiment_counts %>%
  group_by(sentiment) %>%
  summarize(freq=sum(n))

```

```{r}
#| label: sentimentNeg
#| warning: true
neg <- bing_sentiment_counts %>%
  filter(sentiment == "negative")
set.seed(1234) # for reproducibility 
wordcloud::wordcloud(words = neg$word, 
                     freq = neg$n, min.freq = 1,           
                     max.words=200, 
                     random.order=FALSE, 
                     rot.per=0.35,            
                     colors=brewer.pal(8, "Dark2"))
```

The second wordcloud shows words with positive sentiments. Remember the size of the words is relative to the count of all positive words.

```{r}
#| label: sentimentPos
#| warning: true
bing_sentiment_counts <- reduced_df %>%
  dplyr::inner_join(tidytext::get_sentiments("bing")) %>% # only keeps those with sentiments
  #count(linenumber, sentiment, sort = TRUE) %>%
  dplyr::ungroup()

pos <- bing_sentiment_counts %>%
  dplyr::filter(sentiment == "positive")

set.seed(1234) # for reproducibility 
wordcloud::wordcloud(words = pos$word, 
                     freq = pos$n, min.freq = 1,           
                     max.words=200, 
                     random.order=FALSE, 
                     rot.per=0.35,            
                     colors=brewer.pal(8, "Dark2"))
```

The final plot shows trends in negative and positive sentiment over time.

```{r}
#| label: sentimentTrend

bing_sentiment_counts_daily_dt <- data.table::as.data.table(bing_sentiment_counts_daily)
plot_daily_sentiment <- bing_sentiment_counts_daily_dt[, .(n = sum(n)),
                                                       keyby = .(ba_created_at_date, sentiment)]

p <- ggplot2::ggplot(plot_daily_sentiment, aes(
  x = as.Date(ba_created_at_date),
  y = n,
  colour = sentiment
)) +
  geom_line() +
  labs(x = "Date",
       y = "N words per day")
#add_sessionDates(p) only works on half-hourly x axis
```

## Session 1 sentiment

Repeat these negative/postive word clouds but just for the first session which was on `r as.Date(session1DateStart)`.

These are just the tweets for the day of the event and the day after...

Positive words...

```{r}
#| label: posCloudSessionOne
#| warning: true

# default cloud = positive
makeSentimentCloud <- function(dt, sentiFilter = "positive"){
  # start with the selected tweets
  # grab the words
  tweets_words <-  dt %>%
    dplyr::select(clean_text) %>%
    tidytext::unnest_tokens(word, clean_text)
  # count the words
  words <- tweets_words %>% count(word, sort=TRUE)
  # remove the stop words
  data(stop_words)
  
  # remove the stop words - they are usually the most frequent
  # and usually the least interesting
  de_stopped <- words %>%
    dplyr::anti_join(stop_words)
  
  bsc <- de_stopped %>%
    dplyr::inner_join(tidytext::get_sentiments("bing")) %>% # only keeps those with sentiments
    #count(linenumber, sentiment, sort = TRUE) %>%
    ungroup()
  
  results <- list()
  results$ft <- table(bsc$sentiment)
  
  results$nt <- bsc %>%
    group_by(sentiment) %>%
    summarize(freq=sum(n))
  
  # select the sentiment to wordCloud
  filtered <- bsc %>%
    dplyr::filter(sentiment == sentiFilter) # why not working?
  
  set.seed(1234) # for reproducibility
  # seems to get rendered immediately, why?
  wordcloud::wordcloud(words = filtered$word, 
                             freq = filtered$n, min.freq = 1,           
                             max.words=200, 
                             random.order=FALSE, 
                             rot.per=0.35,            
                             colors=brewer.pal(8, "Dark2"))
  
  return(results)
}

session1Tweets <- uniq_tweetsDT[ba_created_at_date >= as.Date(session1DateStart) &
                                 ba_created_at_date <= as.Date(session1DateEnd)+1]

# renders plot immediately (not sure why)
res <- makeSentimentCloud(session1Tweets, 
                   sentiFilter = "positive")

knitr::kable(res$ft, 
             caption = "Number of unique words by sentiment") %>%
  kable_styling()

knitr::kable(res$nt, 
             caption = "Number of words") %>%
  kable_styling()

```

Negative words...

```{r}
#| label: negCloudSessionOne
#| warning: true

res <- makeSentimentCloud(session1Tweets, 
                   sentiFilter = "negative")
```

## Session 2 sentiment

Repeat for session 2 which was on `r as.Date(session2DateStart)`.

These are just the tweets for the day of the event and the day after...

Positive words...

```{r}
#| label: posCloudSession2
#| warning: true
session2Tweets <- uniq_tweetsDT[ba_created_at_date >= as.Date(session2DateStart) &
                                 ba_created_at_date <= as.Date(session2DateEnd)+1]


res <- makeSentimentCloud(session2Tweets, 
                   sentiFilter = "positive")

knitr::kable(res$ft, 
             caption = "Number of unique words by sentiment") %>%
  kable_styling()

knitr::kable(res$nt, 
             caption = "Number of words by sentiment by sentiment") %>%
  kable_styling()
```

Negative words...

```{r}
#| label: negCloudSession2
#| warning: true

res <- makeSentimentCloud(session2Tweets, 
                   sentiFilter = "negative")
```

## Session 3 sentiment

Repeat for session 3 which which was on `r as.Date(session3DateStart)`.

These are just the tweets for the day of the event and the day after...

Positive words...

```{r}
#| label: posCloudSession3
#| warning: true

session3Tweets <- uniq_tweetsDT[ba_created_at_date >= as.Date(session3DateStart) &
                                 ba_created_at_date <= as.Date(session3DateEnd)+1]


res <- makeSentimentCloud(session3Tweets, 
                   sentiFilter = "positive")

knitr::kable(res$ft, 
             caption = "Number of unique words by sentiment") %>%
  kable_styling()

knitr::kable(res$nt, 
             caption = "Number of words by sentiment") %>%
  kable_styling()
```

Negative words

```{r}
#| label: negCloudSession3
#| warning: true

res <- makeSentimentCloud(session3Tweets, 
                   sentiFilter = "negative")
```

## Session 4 sentiment

Repeat for session 4 which was on `r as.Date(session4DateStart)`.

These are just the tweets for the day of the event and the day after...

Positive words...

```{r}
#| label: posCloudSession4
#| warning: true

session4Tweets <- uniq_tweetsDT[ba_created_at_date >= as.Date(session4DateStart) &
                                 ba_created_at_date <= as.Date(session4DateEnd)+1]

res <- makeSentimentCloud(session4Tweets, 
                   sentiFilter = "positive")

knitr::kable(res$ft, 
             caption = "Number of unique words by sentiment") %>%
  kable_styling()

knitr::kable(res$nt, 
             caption = "Number of words by sentiment") %>%
  kable_styling()
```

Negative words...

```{r}
#| label: negCloudSession4
#| warning: true
res <- makeSentimentCloud(session4Tweets, 
                   sentiFilter = "negative")
```

## Session 5 sentiment

Repeat for session 5 which was on `r as.Date(session5DateStart)`.

These are just the tweets for the day of the event and the day after...

Positive words

```{r}
#| label: posCloudSession5
#| warning: true

session5Tweets <- uniq_tweetsDT[ba_created_at_date >= as.Date(session5DateStart) &
                                 ba_created_at_date <= as.Date(session5DateEnd)+1]

res <- makeSentimentCloud(session5Tweets, 
                   sentiFilter = "positive")

knitr::kable(res$ft, 
             caption = "Number of unique words by sentiment") %>%
  kable_styling()

knitr::kable(res$nt, 
             caption = "Number of words by sentiment") %>%
  kable_styling()
```

Negative words...

```{r}
#| label: negCloudSession5
#| warning: true
res <- makeSentimentCloud(session5Tweets, 
                   sentiFilter = "negative")
```

## Session 6 sentiment

Repeat for session 6 which was on `r as.Date(session6DateStart)`.

These are just the tweets for the day of the event and the day after...

Positive words:

```{r}
#| label: posCloudSession6
#| warning: true

session6Tweets <- uniq_tweetsDT[ba_created_at_date >= as.Date(session6DateStart) &
                                 ba_created_at_date <= as.Date(session6DateEnd)+1]

res <- makeSentimentCloud(session6Tweets, 
                   sentiFilter = "positive")

knitr::kable(res$ft, 
             caption = "Number of unique words by sentiment") %>%
  kable_styling()

knitr::kable(res$nt, 
             caption = "Number of words by sentiment") %>%
  kable_styling()
```

Negative words:

```{r}
#| label: negCloudSession6
#| warning: true
res <- makeSentimentCloud(session6Tweets, 
                   sentiFilter = "negative")
```
:::

```{r}
#| label: theEnd
```


## Session 7 sentiment

Repeat for session 7 which was on `r as.Date(session7DateStart)`.

These are just the tweets for the day of the event and the day after...

Positive words

```{r}
#| label: posCloudSession7
#| warning: true

session7Tweets <- uniq_tweetsDT[ba_created_at_date >= as.Date(session7DateStart) &
                                 ba_created_at_date <= as.Date(session7DateEnd)+1]

res <- makeSentimentCloud(session7Tweets, 
                   sentiFilter = "positive")

cap <- paste0("Session 7: ",
              as.Date(session7DateStart),
              " Number of unique words by sentiment"
)
knitr::kable(res$ft, 
             caption = cap)  %>%
  kable_styling()

knitr::kable(res$nt, 
             caption = cap) %>%
  kable_styling()
```

Negative words...

```{r}
#| label: negCloudSession7
#| warning: true
res <- makeSentimentCloud(session6Tweets, 
                   sentiFilter = "negative")
```

:::

```{r}
#| label: theEnd
```
